Robust calibration of numerical models based on relative regret

### 2019-12-18 article.tex
typos and corrections

In this paper, we are going to present how to take into account those uncertainties by defining a notion of robustness based on the distribution of the conditional minimisers and compare it with the minimum in the mean sense, and the minimum of variance.
\end{abstract}

@@

Numerical models are widely used to study or forecast natural phenomenon and improve industrial processes. However, by essence models only partially represents reality and sources of uncertainties are ubiquitous (discretisation errors, missing physical processes, poorly known boundary conditions).
%Many physical phenomena are modelled numerically in order to better understand and/or to predict their behaviour. Because of the complexity and or ignorance of the underlying physics, numerical models are imperfect representations of the reality. 
%
%Various sources of uncertainties can corrupt the model quality, such as the discretisation of the domain, or presence of  small scale phenomena that can not be fully represented. 
Moreover, such uncertainties may be of different nature. 
\cite{walker_defining_2003} proposes to consider two categories of uncertainties. On the one hand aleatoric uncertainties,  coming from the inherent variability of a phenomenon, e.g.\ intrinsic randomness of some environmental variables. On the other hand, epistemic uncertainties coming from a lack of knowledge about the properties and conditions of the phenomena underlying the behaviour of the system under study.
The latter can be accounted for through the introduction of ad-hoc correcting terms in the numerical model, that need to be properly estimated. Thus, reducing the epistemic uncertainty can be done through parameters estimation approaches. 
%when one is having observations of the reality. 
This is usually done using optimal control techniques, leading to an optimisation of a well chosen cost function which is typically built as a comparison with reference observations.
%
An application of such an approach, in the context of ocean circulation modeling, is the estimation of ocean bottom friction parameters in~\cite{das_estimation_1991} and~\cite{boutet_estimation_2015}.

If parameters to be estimated are not the only source of uncertainties, their optimal control 
%However, in the presence of other sources of uncertainties, parameter estimation 
is doomed to overfit the data, \emph{e.g} to artificially introduce errors in the controlled parameter to compensate for other sources. If such uncertainties are of aleatoric nature, then the parameter estimation is only optimal for the observed situation, and may be very poor in other configurations, phenomenon coined as \textit{localized optimization} in~\cite{huyse_free-form_2001}.
%A methodology based on the Hessian matrix has been described in~\cite{ribaud_krigeage_2018-1}.
%\arthur{L\`A AUSSI UNE REF SERAIT BIENVENUE}. 
Taking into account aleatoric uncertainties in optimisation problems takes several names, including \textit{robust optimisation}, \textit{robust design} in~\cite{lelievre_consideration_2016}, or \textit{optimization under uncertainties} \citep{petrone_robustness_2011, seshadri_density-matching_2014,cook_horsetail_2018}.

@@

Let's denote $\mathbf{k} \in \Kspace$ the parameter set to be estimated, to reduce epistemic errors. The aleatoric uncertainties are modelled as a random vector $\mathbf{U}$ whose sample space is $\Uspace$. The probability measure of $\mathbf{U}$ is $\Prob_{\mathbf{U}}$, and its density, if it exists, is $p_{\mathbf{U}}$. %Furthermore, $\mathbf{u}^t$ (used to generate $\mathbf{y}^{\mathrm{obs}}$) is sampled from $\mathbf{U}$.
%
The cost function $J(\mathbf{k}, \mathbf{U})$ is a random variable in this context. %, quantity of interest summarising what we aim to represent with our numerical system $J(\mathbf{k}, \mathbf{U})$ is a random variable that will be further designed as the cost function. 
It is most often defined as the squared norm of a given function $\mathcal{G}(\mathbf{k}, \mathbf{U})$% involving the numerical model. 
\begin{align}
  \label{eq:def_cost_fun}
  J(\mathbf{k},\mathbf{U}) & = \frac12\|\mathcal{G}(\mathbf{k}, \mathbf{U})\|^2
  \end{align}
%$J$ is possibly complemented by a regularisation function that is omitted here for the sake of clarity. 
For instance, in data assimilation, $J$ describes a distance between the output of the numerical model and given observed data, plus generally some regularization terms.
%Calibrating the numerical models consists in the computation of $\kest$ verifying
%\begin{equation}
%  \label{eq:khat_def}
%  \kest = \argmin_{\mathbf{k} \in \Kspace} J(\mathbf{k}, \mathbf{u})
%\end{equation}
%\elise{la il y a un soucis car J(k,U) est une variable aleatoire donc on sait pas ce que ca veut dire de prendre le min} \victor{ici ça serait mieux de mettre un $\mathbf{u}^b$}
%
%where $J$ is then an objective function to minimise,
 %
%For a given $\mathbf{k}$, $J(\mathbf{k},\mathbf{U})$ is a random variable. The main point is to be able to compute a value $\mathbf{k}$ that ``minimises'' in a certain sense the random variable $J(\mathbf{k},\mathbf{U})$. \elise{a revoir. un v.a. ne se minimise pas} 
%
For a practical purpose, we assume that $\forall~ \mathbf{u}$ a realisation of $\mathbf{U}$, finding ${\mathbf{k}}_{\mathbf{u}}$ that minimises the cost function $J(\mathbf{k}, \mathbf{U}=\mathbf{u})$ is a well-posed problem, and that the optimum is unique.
Additionally, the following assumption are made: the cost function is strictly positive, and $\forall~  \mathbf{k} \in \Kspace$, the random variable $J(\mathbf{k},\mathbf{U})$ has finite first- and second-order moments.

@@

As mentioned before, robustness can be understood as satisfactory for a broad range of $\mathbf{u}$, and/or as insensitive as possible to uncertainties encompassed in $\mathbf{U}$. Under this definition, one may design classical robust estimators, either by optimising the moments of the cost function; or based on the distribution of its minimisers.

@@

A lot of different methods are readily available to solve these
minimisation problems. For instance, stochastic Sample Approximation
\citep{juditsky_stochastic_2009,kim_guide_2015} is based on a finite
and fixed set of samples $\{ \mathbf{u}^{i}\}_{i=1 \dots N}$ of
$\mathbf{U}$.
%the sampling of a certain amount of $\mathbf{U}$ 
The estimations at a given $\mathbf{k}$ are computed using standard Monte Carlo, resulting in the following optimisation problems:
\begin{equation}
\hatkmean = \argmin_{\mathbf{k} \in \Kspace} \sum_{i=1}^N  J(\mathbf{k},\mathbf{u}^i)
\end{equation}
and
\begin{equation}
\hatkvar = \argmin_{\mathbf{k} \in \Kspace} \sum_{i=1}^N  \left( J(\mathbf{k},\mathbf{u}^i) -  \sum_{j=1}^N  J(\mathbf{k},\mathbf{u}^j) \right )^2
\end{equation}
%In the following, the hat notation will indicate an estimator. 
If computationally affordable, one can perform these estimations on a regular grid on $\Kspace \times \Uspace$. In case of expensive computer code, one can build a meta model to ease the minimisation, such as Gaussian processes \citep{janusevskis_simultaneous_2010}.
Even though $\kmean$ is a reasonable choice, there is no guarantee that $J(\kmean, \mathbf{U}=\mathbf{u})$ will not reach catastrophic level for some  $\mathbf{u}$. On the other hand, using $\kvar$ will ensure stability of the cost function, but whithout any control of its level of quality.
%
%For a single $\mathbf{k}$, the value of the cost function is controlled in a probabilistic sense by the mean and variance.
%Chebyshev's inequality gives an upper bound on the probability of being more than $\lambda>0$ units from the mean:
%\begin{equation}
%  \label{eq:cheby_ineq}
%  \Prob_{\mathbf{U}}\left[|J(\mathbf{k}% _{\Ex}
%    ,\mathbf{U}) - \mu(\mathbf{k}% _{\Ex}
%    )|> \lambda\right] \leq \frac{\sigma^2(\mathbf{k}% _{\Ex}
%    )}{\lambda^2}
%\end{equation}
%
Ideally, one would want to have a small mean value, and a small variance as well. Multi-objective optimisation is a proper tool to deal with these simultaneous and sometimes concurrent objectives, for exemple by computing the Pareto front of $\left(\mu(\mathbf{k}),\sigma^2(\mathbf{k})\right)$ as done in~\cite{baudoui_optimisation_2012}.

@@

% ----------------------------------------------------------------------------
\subsection{Most probable estimate}
Let us consider the minimal attainable cost in each configuration brought by $\mathbf{u}$.
The resulting conditional minimum is denoted as $J^*$:
\begin{equation}
\label{eq:def_Jstar}
J^*: \mathbf{u} \in\Uspace \longmapsto J^*(\mathbf{u}) = \min_{\mathbf{k}\in \Kspace} J(\mathbf{k},\mathbf{U}=\mathbf{u})
\end{equation}
Similarly, the function of conditional minimisers can then defined by:
  \begin{equation}
  \label{eq:conditional_minimiser}
   \mathbf{k}^*: \mathbf{u}\in\Uspace \longmapsto \mathbf{k}^*(\mathbf{u}) = \mathbf{k}_{\mathbf{u}}=  \argmin_{{\mathbf{k}}\in\Kspace} J({\mathbf{k}},\mathbf{U}=\mathbf{u})
  \end{equation}
Using this function, we can define the corresponding random variable $\mathbf{K}^*$ as
  \begin{equation}
    \label{eq:def_study_minimisers}
    \mathbf{K}^*= \mathbf{k}^*(\mathbf{U}),
  \end{equation}
and its associated density function $p_{\mathbf{K}^*}(\mathbf{k})$, that will be further referred as the density of minimisers.
The mode of this density is called the Most Probable Estimate (MPE) and is noted  $\kmpe$:
\begin{equation}
  \label{eq:MPE}
  \kmpe = \argmax_{\mathbf{k} \in\Kspace} p_{\mathbf{K}^*}(\mathbf{k}) 
\end{equation}
To give some intuition on this estimate, let us imagine that the distribution of minimisers is a dirac centered on $\kmpe$. Then it would mean that this estimate is the minimiser of the cost function whatever the realisation of the uncertain variable, therefore optimal in all conditions. 
If the distribution $p_{\mathbf{K}^*}$ is heavily dominated by a single value, the MPE may be a good candidate for robust control. This is not so obvious in case of a multimodal distribution. 
In general, an analytical form of $p_{\mathbf{K}^*}$ is usually impossible to obtain, so an estimation $\hat{p}_{\mathbf{K}^*}$ must be used, and its maximum computed to get the MPE.

Once again, a set of samples  $\{\mathbf{u}^i\}_{i=1\dots N}$ can be used to compute the set $\{\mathbf{k}_{\mathbf{u}^i}\}_{i=1\dots N}$, from which one can 
approximate  $p_{\mathbf{K}^*}$. The resulting approximation and therefore its mode, is sensitive to the density estimation method. Main methods are KDE  (Kernel Density Estimation)~\cite{elise}, and  EM (Expectation-Maximisation)~\citep{dempster_maximum_1977}.

@@

Despite the fact that those methods are well established, using them in a plug-in approach has some flaws. One of the basic assumption of density estimation is to assume that $\mathbf{K}^*$ is a continuous random variable, hypothesis that may be violated. Worse, the notion of mode is not well defined when the distribution of the minimisers is a discrete-continuous mixture. This may result in inconsistent 
estimations of $\hatkmpe$ when using different methods as illustrated in next subsection.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Numerical illustration}
\label{ssec:num_illu}
Before going further in the explanation of our approach, let us illustrate the nature of previously detailed estimators, $\hatkmean, \hatkvar$ and $\hatkmpe$ on two analytical cost functions. These functions are based on the Branin-Hoo's function, slightly modified to ensure strict positivity:
\begin{align}
  \label{eq:branin_hoo}
 \mathrm{BH}(x_1,x_2) &= \frac{1}{51.95}\left[\left(\bar{x}_2 - \frac{5.1 \bar{x}_1^2}{4\pi^2} + \frac{5\bar{x}_1}{\pi} -6 \right)^2 + \left(10 - \frac{10}{8\pi}\right)\cos(\bar{x}_1)-44.81\right]+2\\
&\text{with } \bar{x}_1 = 3x_1-5,\quad \bar{x}_2 = 3x_2
\end{align}
%
Using Eq.~\eqref{eq:branin_hoo}, we define the two cost functions on $\Kspace \times \Uspace = [0,5] \times [0,5]$ as:
\begin{align*} J_{\mathrm{BH}} : (\mathbf{k},\mathbf{u}) \mapsto \mathrm{BH}(\mathbf{k},\mathbf{u}) \\
J_{\mathrm{BHswap}} : (\mathbf{k},\mathbf{u}) \mapsto \mathrm{BH}(\mathbf{u},\mathbf{k})
\end{align*}
%We will continue to use $J$ as a generic notation for the cost function, when its use is not specific to the Branin-Hoo function.
Even though the functions are quite similar, the asymmetric roles of $\mathbf{k}$ and $\mathbf{u}$ cause different behaviour.

\begin{figure*}[!ht]
  \centering
  \includegraphics[width=12cm]{Figures/branin_side_moments_noscale.pdf}
  \caption{The left column concerns the $J_{\mathrm{BH}}$ function, while the right one deals with $J_{\mathrm{BHswap}}$. Contours of both functions are plotted on the top, and curves of  $\mu(\mathbf{k})$ and $\sigma(\mathbf{k})$ are shown on the bottom (respective scales are note represented). Estimates $\hatkmean$ and $\hatkvar$ are plotted with the dashed line. } \label{fig:branin_moments}
\end{figure*}
%\clearpage % Pour lisibilité dans le preview

The random variable $\mathbf{U}$ is assumed to be uniformly distributed over $\Uspace$.
%The definition and the estimation of $\kmean$ and $\kvar$ is pretty standard.  
The estimations are %$\mathbf{k}\mapsto \mu(\mathbf{k})$ and $\mathbf{k}\mapsto\sigma(\mathbf{k})$ 
based on a $1000 \times 1000$ regular grid over $\Kspace\times\Uspace$. Both cost functions are shown on~Figure~\ref{fig:branin_moments}, top row.

 The left, respectively right, column stands for $J_{\mathrm{BH}}$, respectively $J_{\mathrm{BHswap}}$ function. Functions $\mu(\mathbf{k})$ and $\sigma(\mathbf{k})$ are drawn on the bottom row, respectively in purple and green. The corresponding minimizers $\hatkmean$ and $\hatkvar$ are also plotted.
On this figure, we can observe that $\hatkmean$ and $\hatkvar$ are close for the $J_{\mathrm{BH}}$ function, while being significantly different for the $J_{\mathrm{BHswap}}$ function.
%\clearpage

Similarly $\hatkmpe$ estimations are depicted on Figure~\ref{fig:contours}.  Top row shows the contour plot of both functions as well as the set of conditional minimisers $\{\mathbf{k}_{\mathbf{u}^i}\}_{1\leq i \leq N}$ in red, as defined in Eq.~\eqref{eq:conditional_minimiser}.
The bottom rows presents three approximations of the density of minimisers: the histogram in grey (bin size selected using Freedman-Diaconis from~\cite{freedman_histogram_1981}), the result of a kernel density estimation (KDE) with Gaussian kernels in red (using Scott's rule from~\cite{scott_optimal_1979} for bandwidth selection), and the estimation by a Gaussian mixture. The latter has been calculated with the EM algorithm. The number of Gaussians has been fixed to 3, a guess based on the general shape of the histogram. Respective estimations of $\hatkmpe$ are also depicted using dashed lines.

%In this paper, we used two different methods to estimate the distribution of the minimisers: KDE (kernel density estimation)~\citep{parzen_estimation_1962,bishop_pattern_2006} and the EM algorithm~\citep{dempster_maximum_1977,borman_expectation_2004} to fit a mixture of Gaussian distributions. This algorithm needing the number of densities to fit, we chose to fit three Gaussian, a guess based on the general shape of the histograms.
%Figure~\ref{fig:contours} shows the contour plot, conditional mean and conditional minimisers of the $J_{\mathrm{BH}}$ function (left) and the $J_{\mathrm{BHswap}}$ function (right).

\begin{figure*}
  \centering
  \includegraphics[width=12cm]{Figures/branin_side_66.pdf}
  \caption{Top Left:  $J_{\mathrm{BH}}$ function along with conditional minimisers in red. Bottom left: Estimated density using KDE and EM algorithm and conditional mean. The dashed lines indicate the MPE found using those methods. Right: Same quantities with $J_{\mathrm{BHswap}}$.}
\label{fig:contours}
\end{figure*}

For the $J_{\mathrm{BHswap}}$ function, we can observe that those three methods give consistent results, as $\hatkmpe {}_{,\mathrm{KDE}} =\hatkmpe {}_{,\mathrm{EM}}=\hatkmpe {}_{,\mathrm{histogram}}\approx 0.8$. This is not the case for the  $J_{\mathrm{BH}}$ function however. Using Kernel density estimation (Gaussian), the estimation of $\kmpe$ is $\hatkmpe {}_{,\mathrm{KDE}} \approx 1.5$, while using the histogram and Gaussian mixture, $\hatkmpe {}_{,\mathrm{histogram}}=\hatkmpe {}_{,\mathrm{EM}}=3.8$.
This difference is explained by the accumulation of minimisers at this point: this challenges the assumption that $\mathbf{K}^*$ is continuous. As the density estimation techniques traditionally assume this continuity, the EM algorithm fits this using a normal distribution with a very small variance, while the KDE considers a sum of Gaussian kernels of constant bandwidth, located at the same point. This particular problem highlights an issue with $\kmpe$, as its estimation is possibly sensitive to the density approximation procedure.
% Alternatively, when considering the mixing coefficients of the EM estimation, the Gaussian centered at $1.5$ has the largest weight, of $0.608$. In other words, according to this estimated density, the optimal value $K^*$ has a probability of $0.608$ of being distributed normally around $1.5$.
% Alternatively, we can consider the mixing coefficients of the EM
% density estimation. The Gaussian centered at $1.5$ has been attributed
% a weight of $0.608$, while the one centered at $3.8$ has a mixing
% weight of $0.226$. Ultimately, an optimiser would be more likely to be
% close to $1.5$, but we do not have information on its performance in
% this region around this value. \arthur{je ne comprends pas ce passage}

% as we can argue that both values may be a sensible choice for the MPE. This comes from the fact that the hypothesis that $\mathbf{K}^*$ is a continuous r.v.\ is

Instead of just considering the optimal minimisers, we introduce a bit of leeway, and look for ``acceptably not optimal'' parameters.
 This slackness takes the form of a relaxation coefficient, whose choice leads to a new family of robust estimators, along with a measure of their robustness via the relaxation coefficient.
%This leads us to the definition of a new family of estimators, for which one can attribute a {\it level of robustness} \elise{je me rend bien compte que cette phrase est un peu bizarre ...mais bon c'est l'idee}

%, by transforming the equality of Eq.~\eqref{eq:def_R1} into a inequality.

@@

% We are now going to compute and compare those estimators on the $J_{\mathrm{BH}}$and $J_{\mathrm{BHswap}}$ function.

>>>>

In this paper, we are going to present how to take into account those uncertainties by defining a family of calibration objectives based on the notion of relative-regret with respect to the best attainable performance given the uncertainties and compare it with the minimum in the mean sense, and the minimum of variance.
\end{abstract}

@@

Numerical models are widely used to study or forecast natural phenomena and improve industrial processes. However, by essence models only partially represent reality and sources of uncertainties are ubiquitous (discretisation errors, missing physical processes, poorly known boundary conditions).
%Many physical phenomena are modelled numerically in order to better understand and/or to predict their behaviour. Because of the complexity and or ignorance of the underlying physics, numerical models are imperfect representations of the reality. 
%
%Various sources of uncertainties can corrupt the model quality, such as the discretisation of the domain, or presence of  small scale phenomena that can not be fully represented. 
Moreover, such uncertainties may be of different nature. 
\cite{walker_defining_2003} proposes to consider two categories of uncertainties. On the one hand aleatoric uncertainties,  coming from the inherent variability of a phenomenon, e.g.\ intrinsic randomness of some environmental variables. On the other hand, epistemic uncertainties coming from a lack of knowledge about the properties and conditions of the phenomenon underlying the behaviour of the system under study.
The latter can be accounted for through the introduction of ad-hoc correcting terms in the numerical model, that need to be properly estimated. Thus, reducing the epistemic uncertainty can be done through parameters estimation approaches. 
%when one is having observations of the reality. 
This is usually done using optimal control techniques, leading to an optimisation of a well chosen cost function which is typically built as a comparison with reference observations.
%
An application of such an approach, in the context of ocean circulation modeling, is the estimation of ocean bottom friction parameters in~\cite{das_estimation_1991} and~\cite{boutet_estimation_2015}.

If parameters to be estimated are not the only source of uncertainties, their optimal control 
%However, in the presence of other sources of uncertainties, parameter estimation 
is doomed to overfit the data, \emph{e.g} to artificially introduce errors in the controlled parameter to compensate for other sources. If such uncertainties are of aleatoric nature, then the parameter estimation is only optimal for the observed situation, and may be very poor in other configurations, phenomenon coined as \textit{localized optimisation} in~\cite{huyse_free-form_2001}.
%A methodology based on the Hessian matrix has been described in~\cite{ribaud_krigeage_2018-1}.
%\arthur{L\`A AUSSI UNE REF SERAIT BIENVENUE}. 
Taking into account aleatoric uncertainties in optimisation problems takes several names, including \textit{robust optimisation}, \textit{robust design} in~\cite{lelievre_consideration_2016}, or \textit{optimisation under uncertainties} \citep{petrone_robustness_2011, seshadri_density-matching_2014,cook_horsetail_2018}.

@@

Let's denote $\mathbf{k} \in \Kspace$ the parameter set to be estimated to reduce epistemic errors. The aleatoric uncertainties are modelled as a random vector $\mathbf{U}$ whose sample space is $\Uspace$. The probability measure of $\mathbf{U}$ is $\Prob_{\mathbf{U}}$, and its density, if it exists, is $p_{\mathbf{U}}$. %Furthermore, $\mathbf{u}^t$ (used to generate $\mathbf{y}^{\mathrm{obs}}$) is sampled from $\mathbf{U}$.
%
The cost function $J(\mathbf{k}, \mathbf{U})$ is a random variable in this context. %, quantity of interest summarising what we aim to represent with our numerical system $J(\mathbf{k}, \mathbf{U})$ is a random variable that will be further designed as the cost function. 
It is most often defined as the squared norm of a given function $\mathcal{G}(\mathbf{k}, \mathbf{U})$% involving the numerical model. 
\begin{align}
  \label{eq:def_cost_fun}
  J(\mathbf{k},\mathbf{U}) & = \frac12\|\mathcal{G}(\mathbf{k}, \mathbf{U})\|^2
  \end{align}
%$J$ is possibly complemented by a regularisation function that is omitted here for the sake of clarity. 
For instance, in data assimilation, $J$ describes a distance between the output of the numerical model and given observed data, plus generally some regularization terms.
%Calibrating the numerical models consists in the computation of $\kest$ verifying
%\begin{equation}
%  \label{eq:khat_def}
%  \kest = \argmin_{\mathbf{k} \in \Kspace} J(\mathbf{k}, \mathbf{u})
%\end{equation}
%\elise{la il y a un soucis car J(k,U) est une variable aleatoire donc on sait pas ce que ca veut dire de prendre le min} \victor{ici ça serait mieux de mettre un $\mathbf{u}^b$}
%
%where $J$ is then an objective function to minimise,
 %
%For a given $\mathbf{k}$, $J(\mathbf{k},\mathbf{U})$ is a random variable. The main point is to be able to compute a value $\mathbf{k}$ that ``minimises'' in a certain sense the random variable $J(\mathbf{k},\mathbf{U})$. \elise{a revoir. un v.a. ne se minimise pas} 
%
For a practical purpose, we assume that for every realisation $\mathbf{u}\in\Uspace$ of $\mathbf{U}$, % $\forall~ \mathbf{u}$ a realisation of $\mathbf{U}$
finding ${\mathbf{k}}_{\mathbf{u}}\in \Kspace$ that minimises the cost function $\mathbf{k}\mapsto J(\mathbf{k}, \mathbf{U}=\mathbf{u})$ is a well-posed problem, and that the optimum is unique.
Additionally, the following assumptions are made: the cost function is strictly positive, and $\forall~  \mathbf{k} \in \Kspace$, the random variable $J(\mathbf{k},\mathbf{U})$ has finite first- and second-order moments.

@@

As mentioned before, robustness can be understood as satisfactory for a broad range of $\mathbf{u}$, and/or as insensitive as possible to uncertainties encompassed in $\mathbf{U}$. Under this definition, one may design robust estimators, either by using the moments of the cost function; or by exploiting the distribution of its minimisers.

@@

A lot of different methods are readily available to solve these
minimisation problems. For instance, stochastic Sample Approximation
\citep{juditsky_stochastic_2009,kim_guide_2015} is based on a finite
and fixed set of samples $\{ \mathbf{u}^{i}\}_{i=1 \dots N}$ of
$\mathbf{U}$.
%the sampling of a certain amount of $\mathbf{U}$ 
The estimations at a given $\mathbf{k}$ are computed using standard Monte Carlo, resulting in the following optimisation problems:
\begin{equation}
\hatkmean = \argmin_{\mathbf{k} \in \Kspace} \sum_{i=1}^N  J(\mathbf{k},\mathbf{u}^i)
\end{equation}
and
\begin{equation}
\hatkvar = \argmin_{\mathbf{k} \in \Kspace} \sum_{i=1}^N  \left( J(\mathbf{k},\mathbf{u}^i) -  \sum_{j=1}^N  J(\mathbf{k},\mathbf{u}^j) \right )^2
\end{equation}
%In the following, the hat notation will indicate an estimator. 
If computationally affordable, one can perform these estimations on a regular grid on $\Kspace \times \Uspace$. In case of expensive computer code, one can build a meta model to ease the minimisation, such as Gaussian processes \citep{janusevskis_simultaneous_2010}.
Even though $\kmean$ is a reasonable choice, there is no guarantee that $J(\kmean, \mathbf{U}=\mathbf{u})$ will not reach catastrophic level for some  $\mathbf{u}$. On the other hand, using $\kvar$ will ensure stability of the cost function, but without any control of its performance. %level of quality.
%
%For a single $\mathbf{k}$, the value of the cost function is controlled in a probabilistic sense by the mean and variance.
%Chebyshev's inequality gives an upper bound on the probability of being more than $\lambda>0$ units from the mean:
%\begin{equation}
%  \label{eq:cheby_ineq}
%  \Prob_{\mathbf{U}}\left[|J(\mathbf{k}% _{\Ex}
%    ,\mathbf{U}) - \mu(\mathbf{k}% _{\Ex}
%    )|> \lambda\right] \leq \frac{\sigma^2(\mathbf{k}% _{\Ex}
%    )}{\lambda^2}
%\end{equation}
%
Ideally, one would want to have a small expectation and a small variance at the same time. Multi-objective optimisation is a proper tool to deal with these simultaneous and sometimes concurrent objectives, for exemple by computing the Pareto front of $\left(\mu(\mathbf{k}),\sigma^2(\mathbf{k})\right)$ as done in~\cite{baudoui_optimisation_2012}.

@@

% ----------------------------------------------------------------------------
\subsection{Most probable estimate}
Let us consider the minimal cost attainable in each configuration brought by $\mathbf{u}$.
The resulting conditional minimum is denoted $J^*$:
\begin{equation}
\label{eq:def_Jstar}
J^*: \mathbf{u} \in\Uspace \longmapsto J^*(\mathbf{u}) = \min_{\mathbf{k}\in \Kspace} J(\mathbf{k},\mathbf{U}=\mathbf{u})
\end{equation}
Similarly, the function of conditional minimisers can then defined by:
  \begin{equation}
  \label{eq:conditional_minimiser}
   \mathbf{k}^*: \mathbf{u}\in\Uspace \longmapsto \mathbf{k}^*(\mathbf{u}) = \mathbf{k}_{\mathbf{u}}=  \argmin_{{\mathbf{k}}\in\Kspace} J({\mathbf{k}},\mathbf{U}=\mathbf{u})
  \end{equation}
Using this function, we can define the corresponding random variable $\mathbf{K}^*$ as
  \begin{equation}
    \label{eq:def_study_minimisers}
    \mathbf{K}^*= \mathbf{k}^*(\mathbf{U}),
  \end{equation}
and its associated density function $p_{\mathbf{K}^*}(\mathbf{k})$, that will be further referred as the density of minimisers.
The mode of this density is called the Most Probable Estimate (MPE) and is noted  $\kmpe$:
\begin{equation}
  \label{eq:MPE}
  \kmpe = \argmax_{\mathbf{k} \in\Kspace} p_{\mathbf{K}^*}(\mathbf{k}) 
\end{equation}
To give some intuition on this estimate, let us imagine that the distribution of minimisers is a dirac centered on $\kmpe$. Then it would mean that this estimate is the minimiser of the cost function whatever the realisation of the uncertain variable, therefore optimal in all conditions. 
If the distribution $p_{\mathbf{K}^*}$ is heavily dominated by a single value, the MPE may be a good candidate for robust control. This is not so obvious in case of a multimodal distribution.

In general, an analytical form of $p_{\mathbf{K}^*}$ is usually impossible to obtain, so an estimation $\hat{p}_{\mathbf{K}^*}$ must be used, and its maximum computed to get the MPE.
Once again, a set of samples  $\{\mathbf{u}^i\}_{i=1\dots N}$ can be used to compute the set $\{\mathbf{k}_{\mathbf{u}^i}\}_{i=1\dots N}$, from which one can 
approximate  $p_{\mathbf{K}^*}$. The resulting approximation and therefore its mode, is sensitive to the density estimation method. Main methods are KDE  (Kernel Density Estimation)~\cite{elise}, and  EM (Expectation-Maximisation)~\citep{dempster_maximum_1977}.

@@

In practice, despite the fact that those methods are well established, using them in a plug-in approach has some flaws. One of the basic assumption of density estimation is to assume that $\mathbf{K}^*$ is a continuous random variable, hypothesis that may be violated. Worse, the notion of mode is not well defined when the distribution of the minimisers is a discrete-continuous mixture. This may result in inconsistent 
estimations of $\hatkmpe$ when using different methods as illustrated in next subsection.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Numerical illustration}
\label{ssec:num_illu}
Before going further in the explanation of our approach, let us illustrate the nature of previously detailed estimators, $\hatkmean, \hatkvar$ and $\hatkmpe$ on two analytical cost functions. These functions are based on the Branin-Hoo's function, slightly modified to ensure strict positivity:
\begin{align}
  \label{eq:branin_hoo}
 \mathrm{BH}(x_1,x_2) &= \frac{1}{51.95}\left[\left(\bar{x}_2 - \frac{5.1 \bar{x}_1^2}{4\pi^2} + \frac{5\bar{x}_1}{\pi} -6 \right)^2 + \left(10 - \frac{10}{8\pi}\right)\cos(\bar{x}_1)-44.81\right]+2\\
&\text{with } \bar{x}_1 = 3x_1-5,\quad \bar{x}_2 = 3x_2
\end{align}
%
Using Eq.~\eqref{eq:branin_hoo}, we define the two cost functions on $\Kspace \times \Uspace = [0,5] \times [0,5]$ as:
\begin{align*} \JBH : (\mathbf{k},\mathbf{u}) \mapsto \mathrm{BH}(\mathbf{k},\mathbf{u}) \\
\JBHS : (\mathbf{k},\mathbf{u}) \mapsto \mathrm{BH}(\mathbf{u},\mathbf{k})
\end{align*}
%We will continue to use $J$ as a generic notation for the cost function, when its use is not specific to the Branin-Hoo function.
Even though the functions are quite similar, the asymmetric roles of $\mathbf{k}$ and $\mathbf{u}$ cause different behaviour in the estimations.

\begin{figure*}[!ht]
  \centering
  \includegraphics[width=10cm]{Figures/branin_side_moments_noscale.pdf}
  \caption{The left column concerns $\JBH$, while the right one deals with $\JBHS$. Contours of both functions are plotted on the top, and curves of  $\mu(\mathbf{k})$ and $\sigma(\mathbf{k})$ are shown on the bottom (respective scales are not displayed). Estimates $\hatkmean$ and $\hatkvar$ are plotted with the dashed line.}
\label{fig:branin_moments}
\end{figure*}
%\clearpage % Pour lisibilité dans le preview

The random variable $\mathbf{U}$ is assumed to be uniformly distributed over $\Uspace$.
%The definition and the estimation of $\kmean$ and $\kvar$ is pretty standard.  
The estimations are %$\mathbf{k}\mapsto \mu(\mathbf{k})$ and $\mathbf{k}\mapsto\sigma(\mathbf{k})$ 
based on a $1000 \times 1000$ regular grid over $\Kspace\times\Uspace$. Both cost functions are shown on the top of~Figure~\ref{fig:branin_moments}.

The left, (resp.\ right), column stands for $\JBH$ (resp.\ $\JBHS$).
Functions $\mu(\mathbf{k})$ and $\sigma(\mathbf{k})$ are drawn on the bottom row, respectively in purple and green. The corresponding minimisers $\hatkmean$ and $\hatkvar$ are also plotted.
On this figure, we can observe that $\hatkmean$ and $\hatkvar$ are close for $\JBH$, while being significantly different for $\JBHS$.
%\clearpage

Similarly, estimations of $\hatkmpe$ are depicted on Figure~\ref{fig:contours}.The top row shows the contours of both functions as well as the set of conditional minimisers $\{\mathbf{k}_{\mathbf{u}^i}\}_{1\leq i \leq N}$ in red, as defined in Eq.~\eqref{eq:conditional_minimiser}.
The bottom row presents three approximations of the density of minimisers: the histogram in grey (bin size selected using Freedman-Diaconis from~\cite{freedman_histogram_1981}), the result of a kernel density estimation (KDE) with Gaussian kernels in red (using Scott's rule from~\cite{scott_optimal_1979} for bandwidth selection), and the estimation by a Gaussian mixture, calculated with the EM algorithm. The number of Gaussians has been fixed to 3, a guess based on the general shape of the histogram. Respective estimations of $\hatkmpe$ are also depicted using dashed lines. 
\begin{figure}[!h]
  \centering
  \includegraphics[width=10cm]{Figures/branin_side_66.pdf}
  \caption{Top Left:  $\JBH$ along with conditional minimisers in red. Bottom left: Estimated densities using KDE and EM algorithm and the histogram. The dashed lines indicate the MPE found using those methods. Right: Same quantities with $\JBHS$.}
\label{fig:contours}
\end{figure}

For $\JBHS$, we can observe that those three methods give consistent results, as $\hatkmpe {}_{,\mathrm{KDE}} =\hatkmpe {}_{,\mathrm{EM}}=\hatkmpe {}_{,\mathrm{histogram}}\approx 0.8$. This is not the case for $\JBH$: using Kernel density estimation (Gaussian), the estimation of $\kmpe$ is $\hatkmpe {}_{,\mathrm{KDE}} \approx 1.5$, while using the histogram and Gaussian mixture, $\hatkmpe {}_{,\mathrm{histogram}}=\hatkmpe {}_{,\mathrm{EM}}=3.8$.

This difference is explained by the accumulation of minimisers at this point: this challenges the assumption that $\mathbf{K}^*$ is continuous. As the density estimation techniques traditionally assume this continuity, the EM algorithm fits this using a normal distribution with a very small variance, while the KDE considers a sum of Gaussian kernels of constant bandwidth, located at the same point. This particular problem highlights an issue with $\kmpe$, as its estimation is possibly sensitive to the density approximation procedure.

Instead of just considering the optimal minimisers, we introduce a bit of leeway, and look for ``acceptably not optimal'' parameters.
 This slackness takes the form of a relaxation coefficient and its choice defines a new family of robust estimators, where each one of its member carries information on its robustness through this coefficient.
%This leads us to the definition of a new family of estimators, for which one can attribute a {\it level of robustness} \elise{je me rend bien compte que cette phrase est un peu bizarre ...mais bon c'est l'idee}

@@

% We are now going to compute and compare those estimators on the $\JBH$and $\JBHS$ function.

% ----------------------------------------------------------------------------
\subsection{Numerical Illustration}
\label{ssec:numerical_illustration}
% Numerical application}

In this Section, we will compare the different estimators introduced previously and summarised in Table~\ref{tab:RO_recap} on $\JBH$ and $\JBHS$.
